paper

### 2019-01-10 Report/paper.tex
df_replaced
add changes

    \documentclass[journal,twocolumn]{IEEEtran}
\usepackage[noadjust]{cite}

@@

\begin{abstract}
    In this report we explore ways to automatically analyze ICLS and CSCL
papers to gain an understanding of the community formed by conferences
participants and of the research being conducted within the community.
After describing how to extract data from the original dataset we we
derive insights into the community using social network analysis and
{[}Nour: NLP etc{]}. We discover patterns of collaboration between
participants, their migration and the global spread of the community.
{[}Something about the insights in keywords/NLP/Co-Citation{]}
\end{abstract}
    \hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

@@

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_30_0.png}\end{center}
        \caption{}
        \label{}
    \end{figure}

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_32_0.png}\end{center}
        \caption{}
        \label{}
    \end{figure}

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_34_0.png}\end{center}
        \caption{}
        \label{}
    \end{figure}

@@

Our first graph is constructed using the references of each document. It
is constructed as follow, for each two documents, if there is an
intersection between the references of both documents then there is a
link. The weights of the edges are the length of the intersection set.
The intuition behind is that the more the weight is high, the more the
documents are likely to be about the same subjects since they are citing
the same documents. From this graph, we create a subgraph of it, we
filter the edges keeping only those who have a weight strictly higher
than 3. It means that, each node of the subgraph, is related to at least
one other node with 4 same documents referenced.

@@

    \hypertarget{third-graph-title-to-find}{%
\paragraph{Third graph: ??? Title to
find}\label{third-graph-title-to-find}}

@@

@@

To get insight into collaborations within the community we build graphs
based on co-authorship of papers. To build the graph structure and
visualize it the python package \emph{networkx} was used. First, we
built a co-authorship graph based on individual authors. Each node
represents an author, and two nodes are connected if the authors
collaborated on a paper. Other works \cite{cheong2009social},
\cite{hesford2006management} have used a directed graph that only adds a
link from the first author of the paper to all other authors. However,
as we are more interested in collaboration patterns and less in finding
the most influential first author, we build an undirected graph with
edges between all co-authors. On the downside, yields a very dense graph
that can not directly be interpreted. Hence, we prune the initial graph
by removing nodes that do not satisfy additional conditions. Intuitively
it is clear that a singular collaboration on a paper does not
necessarily indicate that two people even know each other. A first
condition was to only link two nodes together if the authors
corresponding to each node have appear as co-authors in more than one
conference. The graph produced by this approach is much more sparse and
contains much fewer authors. Additionally all connected components
contain fewer than 50 authors, making it possible to consider each
author individually. Such a component is shown in figure \ref{fig_plot}.
A second alternative approach was to consider the strength of bounds. In
the graph authors are only linked if they have written at least 2 papers
together. This method yields a much larger component --containing 441
authors-- as this second condition is weaker. To get an understanding of
author to author relations in this component we need to further prune
away parts of the graph that are not insightful. When we look at
individual author within the community, we are interested well
collaborative individuals. Hence we remove nodes that have collaborated
with fewer than 3 connections to other authors. The resulting graph can
be seen in figure \ref{author_collab_large}. Details creating such a
plot can be found in the section dedicated to this below.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_48_0.png}\end{center}
        \caption{One of the connected components of the co-authorship graph}
        \label{fig_plot}
    \end{figure}

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_49_0.png}\end{center}
        \caption{Largest Component}
        \label{author_collab_large}
    \end{figure}

    Focusing in on the authors is not the only way to interpret the
co-authorship graph. One should also consider the overall structure of
the graph. As graph is not connected, and contains one large main
component with 1141 authors and 208 very small components with less than
25 authors in them we focus on analyzing the structure of the main
component of the graph. First we aim to detect sub-communities within
the graph. Modularity is one measure used to find such communities. It
measures the connection strength within a community relative to the
strength of outgoing connections. Networks with high modularity have
dense connections between the nodes within modules but sparse
connections between nodes in different modules. A non-parametric
algorithm for community detection is the Louvain Method
\cite{blondel2008fast}. We used the implementation of Thomas Aynaud
\cite{louvain}. This gives a good partitioning of the graph into 26
communities of median size 42.84. This is quite a large number of
communities, so we consider an alternative way to yield bigger
partitions. An alternate way to find subcommunities is to look at
normalized cuts, which partition the nodes into groups with many
within-group connections and relatively few between-group connections.
The advantage of this method is that we can choose how many cuts to
make, and we are thus able to define the size of subcommunities. Finding
normalized cuts in a graph can be done using \emph{sklearns} spectral
clustering algorithm by passing it the adjacency matrix of the graph.
With this method we are able to neatly separate the main component graph
in three large communities. We note that the clusters found using method
one are contained in the clusters found by method 2. See Figure \ref{}
and Figure \ref{}.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_51_0.png}\end{center}
        \caption{}
        \label{}
    \end{figure}

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_52_0.png}\end{center}
        \caption{}
        \label{}
    \end{figure}

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_54_0.png}\end{center}
        \caption{Different formats used to reference institution}
        \label{unigraph}
    \end{figure}

    \textbf{Plotting of co-authorship graphs}: Graphs were visualised with
the built in plot functions and graph layouts of networkx. The force
directed layout was used as it clusters together nodes with strong
connections. Then the size of each node was changed based on the
position of the node in the graph. Central nodes with a lot of
connections should be larger, while nodes with fewer connections should
be smaller. To achieve this we used the measures of degree centrality
for graphs with few nodes and edges and betweenness centrality for
larger graphs . Degree centrality measures the fraction of degrees
(number of connections) a node has with respect to the total number of
connections in the graph. Betweenness centrality of a node is the sum of
the fraction of all-pairs shortest paths that pass through it. Degree
centrality is more adept for smaller graphs as their structure is still
easy to interpret. As by construction betweenness takes into
consideration the degrees of it neighboring nodes, it better structures
the nodes large graphs into central and less central nodes.

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_58_0.png}\end{center}
        \caption{Distribution of papers published over the years}
        \label{fig1}
    \end{figure}

@@

We find that participants from 372 institutions participated over the
last 4 years, with a median of 128 institutions per year. Numbers of
institutions participating are increasing each year, with a noticeable
peak in 2018, which participants from 261 institutions. This is a
considerable up from the last conference ICLS in 2016 which had 135
participating institutions. Most participants are from the US, with a
average of 50\% of all participants being from the US. We can also
observe a trend of increase in US participants with only 35\% of US
participants in 2015 and 51\% in 2017, followed by 64\% in 2018.
Overall, participants from 41 nations have participated over the 4
years, which an average of 28 nations represented each conference. After
the US, participants of Germany, Canada and Singapore make up the
biggest proportion of participants \(\ref{figrep}\).

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Collaboration
\item
  Movement of participants
\item
  Citation
\end{enumerate}

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_60_0.png}\end{center}
        \caption{Different formats used to reference institution}
        \label{figrep}
    \end{figure}

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_62_0.png}\end{center}
        \caption{Collaborative papers between countries}
        \label{figcollab}
    \end{figure}

@@

Using the data we have of participants over the years, and knowing their
affiliation, we can now find members that change institutions over the
years. Over the years of recording we find 14 participants that have
changed moved countries and 47 that have moved institutions. We can see
that most movement occured within the US respectively Europe,
{[}fig\_move{]}.

    \hypertarget{research-topics-in-the-community}{%
\subsection{Research topics in the
Community}\label{research-topics-in-the-community}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Keywords
\item
  Outside influence: Conferences being sited
\end{enumerate}

>>>>

    \documentclass[journal,onecolumn]{IEEEtran}
\usepackage[noadjust]{cite}

@@

\begin{abstract}
    In this report we explore ways to automatically analyze ICLS and CSCL
papers to gain an understanding of the community formed by conference
participants and of the research being conducted within the community.
After describing how to extract data from the original dataset we derive
insights into the community using network analysis, natural language
processing and text mining. We discover patterns of collaboration
between participants, their migration and the global spread of the
community. We extract keywords and find clusters of referenced papers
and documents within the conferences.
\end{abstract}
    \hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_28_0.png}\end{center}
        \caption{}
        \label{}
    \end{figure}

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_31_0.png}\end{center}
        \caption{}
        \label{}
    \end{figure}

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_33_0.png}\end{center}
        \caption{}
        \label{}
    \end{figure}

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_35_0.png}\end{center}
        \caption{}
        \label{}
    \end{figure}

@@

Our first graph is constructed using the references of each document. It
is constructed as follow, for each two documents, if there is an
intersection between the references of both documents then there is a
link. The weights of the edges are the length of the intersection set.
The intuition behind is that the more the weight is high, the more the
documents are likely to be about the same subject since they are citing
the same documents. From this graph, we create a subgraph of it, we
filter the edges keeping only those who have a weight strictly higher
than 3. It means that, each node of the subgraph, is related to at least
one other node with 4 same documents referenced.

\begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}212}]:} <IPython.core.display.HTML object>
\end{Verbatim}

@@

    \hypertarget{third-graph-exploratory-references-graph}{%
\paragraph{Third graph: exploratory references
graph}\label{third-graph-exploratory-references-graph}}

@@

\begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}105}]:} <IPython.core.display.HTML object>
\end{Verbatim}

@@

To get insight into collaborations within the community we build graphs
based on co-authorship of papers. To build the graph structure and
visualize it the python package \emph{networkx} was used. When
constructing the graph, we associate qualitative features --such as the
name and institution of the author that is represented by a node-- to
the nodes and edges of the graph.

\hypertarget{graph-construction}{%
\subsubsection{Graph Construction:}\label{graph-construction}}

First, we built a co-authorship graph based on individual authors. Each
node represents an author, and two nodes are connected if the authors
collaborated on a paper. Other works \cite{cheong2009social},
\cite{hesford2006management} have used a directed graph that only adds a
link from the first author of the paper to all other authors. However,
as we are more interested in collaboration patterns and less in finding
the most influential first author, we build an undirected graph with
edges between all co-authors. On the downside, this yields a very dense
graph. Hence, we prune the initial graph by removing nodes that do not
satisfy additional conditions. Intuitively it is clear that a singular
collaboration on a paper does not necessarily indicate that two people
even know each other. A first condition was to only link two nodes
together if the authors corresponding to each node have appear as
co-authors in more than one conference. The graph produced by this
approach is much more sparse and contains much fewer authors.
Additionally all connected components contain fewer than 50 authors,
making it possible to consider each author individually. Such a
component is shown in figure \ref{fig_plot}.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_50_0.png}\end{center}
        \caption{One of the connected components of the co-authorship graph}
        \label{fig_plot}
    \end{figure}

    A second alternative approach was to consider the strength of bounds. In
the graph authors are only linked if they have written at least 2 papers
together. This method yields a much larger component --containing 441
authors-- as this second condition is weaker. To get an understanding of
author to author relations in this component we need to further prune
away parts of the graph that are not insightful. When we look at
individual author within the community, we are interested well
collaborative individuals. Hence we remove nodes that have collaborated
with fewer than 3 connections to other authors. The resulting graph can
be seen in figure \ref{author_collab_large}. Details creating such a
plot can be found in the section dedicated to this below.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_52_0.png}\end{center}
        \caption{Largest Component}
        \label{author_collab_large}
    \end{figure}

    Focusing in on the authors is not the only way to interpret the
co-authorship graph. One should also consider the overall structure of
the graph. As graph is not connected, and contains one large main
component with 1141 authors and 208 very small components with less than
25 authors in them we focus on analyzing the structure of the main
component of the graph. First we aim to detect sub-communities within
the graph. Modularity is one measure used to find such communities. It
measures the connection strength within a community relative to the
strength of outgoing connections. Networks with high modularity have
dense connections between the nodes within modules but sparse
connections between nodes in different modules. A non-parametric
algorithm for community detection is the Louvain Method
\cite{blondel2008fast}. We used the implementation of Thomas Aynaud
\cite{louvain}. This gives a good partitioning of the graph into 26
communities of median size 42.84. This is quite a large number of
communities, so we consider an alternative way to yield bigger
partitions. An alternate way to find subcommunities is to look at
normalized cuts, which partition the nodes into groups with many
within-group connections and relatively few between-group connections.
The advantage of this method is that we can choose how many cuts to
make, and we are thus able to define the size of subcommunities. Finding
normalized cuts in a graph can be done using \emph{sklearns} spectral
clustering algorithm by passing it the adjacency matrix of the graph.
With this method we are able to neatly separate the main component graph
in three large communities. We note that the clusters found using method
one are contained in the clusters found by method 2. See Figure
\ref{Louvain} and Figure \ref{normalized}.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_54_0.png}\end{center}
        \caption{Subcommunities by Louvain}
        \label{Louvain}
    \end{figure}

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_55_0.png}\end{center}
        \caption{Subcommunities via normalized cuts}
        \label{normalized}
    \end{figure}

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_57_0.png}\end{center}
        \caption{Different formats used to reference institution}
        \label{unigraph}
    \end{figure}

    \hypertarget{numerical-analysis-of-graphs}{%
\subsubsection{Numerical analysis of
graphs:}\label{numerical-analysis-of-graphs}}

To analyze the structure of the graphs themselves we calculated the
following measures:

\begin{itemize}
\tightlist
\item
  Diameter: longest shortest path in the graph.
\item
  Average clustering coefficient: Measure of how much the nodes in a
  graph tend to cluster together. Average over all clustering
  coefficients. The clustering coefficient of a node is the fraction of
  triangles prassing through the node over all possible triangles
  through that node \cite{watts1998collective}.
\item
  Density: the number of edges in the graph divided by the maximum
  number of possible edges for a graph with the same number of nodes
\end{itemize}

    \hypertarget{plotting-of-co-authorship-graphs}{%
\subsubsection{Plotting of co-authorship
graphs:}\label{plotting-of-co-authorship-graphs}}

Graphs were visualised with the built in plot functions and graph
layouts of networkx. The force directed layout was used as it clusters
together nodes with strong connections. Then the size of each node was
changed based on the position of the node in the graph. Central nodes
with a lot of connections should be larger, while nodes with fewer
connections should be smaller. To achieve this we used the measures of
degree centrality for graphs with few nodes and edges and betweenness
centrality for larger graphs . Degree centrality measures the fraction
of degrees (number of connections) a node has with respect to the total
number of connections in the graph. Betweenness centrality of a node is
the sum of the fraction of all-pairs shortest paths that pass through
it. Degree centrality is more adept for smaller graphs as their
structure is still easy to interpret. As by construction betweenness
takes into consideration the degrees of it neighboring nodes, it better
structures the nodes large graphs into central and less central nodes.

@@

In this section we present first insights gathered during the
Bibliographic analysis.

@@

We found that participants from 372 institutions participated over the
last 4 years, with a median of 128 institutions per year. The number of
institutions participating increased each year, with a noticeable peak
in 2018, with participants from 261 institutions. This is a considerable
increase from the last conference ICLS in 2016 which had 135
participating institutions. Most participants are from the US, with a
average of 50\% of all participants being from the US. We could also
observe a trend of increase in US participants with only 35\% of US
participants in 2015 and 51\% in 2017, followed by 64\% in 2018.
Overall, participants from 41 nations have participated over the 4
years, with an average of 28 nations represented at each conference.
After the US, participants of Germany, Canada and Singapore make up the
biggest proportion of participants \ref{fig_rep}.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_64_0.png}\end{center}
        \caption{Different formats used to reference institution}
        \label{figrep}
    \end{figure}

@@

*\textbf{International Collaboration:} The U.S. again stands out as a
major player, with most collaborations happening between the U.S. and
other countries. Out of 851 papers, 153 were international
collaborations. Out of these 153, 108 involved at least one U.S.
participant. Excluding U.S. participation we find figure \ref{}. We
noted that Finland is very collaborative, and that there exists a strong
bond between Sweden and Israel. Additionally we observed that Asian
nations seem to collaborate almost exclusively with each other or
Canada.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_66_0.png}\end{center}
        \caption{Collaborative papers between countries}
        \label{figcollab}
    \end{figure}

@@

Using the data we have of participants over the years, and knowing their
affiliation, we can now find members that change institutions over the
years. Over the years of recording we find 14 participants that have
changed moved countries and 47 that have moved institutions. We can see
that most movement occured within the US respectively Europe,
\ref{fig_move}.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_68_0.png}\end{center}
        \caption{Movement of participants across institutions}
        \label{fig_move}
    \end{figure}

    \hypertarget{clustering-and-referecences-graphs-results}{%
\subsection{Clustering and referecences graphs
results}\label{clustering-and-referecences-graphs-results}}

The distribution of referenced publications years are plotted in the
graph below. We can see from the graph that there is a kind of a trend,
each two, three years there is a pick of documents referenced followed
by a two or three years of stagnancy. We can also notice that
throughouts the year each pick is higher than the older one, which can
be explained by the fact that The International Society of the Learning
Sciences community is investigating more and more tools to facilitate
learning over the years.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_70_0.png}\end{center}
        \caption{}
        \label{}
    \end{figure}

    Since the number of documents is increasing over the years, we also
wanted to investigate more if the relevance of all the documents
published is increasing too. Using Pagerank algorithm and the In-degree
of each nodes to get the most influancial documents referenced of the
citation graph that we constructed (graph3) before. We filtered the
first 100 documents and then contructed a histogram of the counts. We
obtain the figure below:

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_72_0.png}\end{center}
        \caption{}
        \label{}
    \end{figure}

    We can see that in the top 100 cited documents more than 12\% are from
2006, 8\% are from 2013, 6\% from 2014\ldots{} We can not say that we
have the same trend as before, and the relevance of the documents is
increasing in the same way than with the number of publications but what
we can say is that more than 50\% of the most cited documents are from
2006 until 2016, we is 10 years. The other half is between 1977 until
2005 which represents more than 25 years. On top of that, 2016 is the
most important year in term of publications and maybe some documents did
not have enough time to be cited yet.

    The top 10 of the most influencial and most important referenced papers
are the following:

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_75_0.png}\end{center}
        \caption{}
        \label{}
    \end{figure}

    We also noticed from the first graph that some documents were more
likely to cited together. We can say that the more two documents are
likely to cited together the more it is probable that both documents are
about the same subject From the table below we can also notice that some
papers cites like to cite the same authors together, which is logic in a
sense, because if some authors always treat about the same subjects then
they are experts in that field and a lot of their publications are
relevant and can be cited together.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_77_0.png}\end{center}
        \caption{}
        \label{}
    \end{figure}

    Using this graph, we also want to extract if there are some clusters of
papers cited together. As we explained before, we run strongly connected
algorithm to check if some cited documents are more likely to cited
together, and this will give us clusters of cited documents that are
more likely to be about the same subjects. We can visualise one cluster,
the most important one.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_80_0.png}\end{center}
        \caption{}
        \label{}
    \end{figure}

\begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}185}]:} <IPython.core.display.HTML object>
\end{Verbatim}

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_82_0.png}\end{center}
        \caption{}
        \label{}
    \end{figure}

    The largest strongly connected component is the composed of 8 nodes and
56 edges. This documents are all cited together more than 4 times. You
can investingate the edges and the weights in the interactive graph
constructed above. We wanted all the documents to be linked to each
others to be more precise, but if we only look to giant connected
component, we will get a network of 27 referenced paper with 142 egdes.

\begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}236}]:} <IPython.core.display.HTML object>
\end{Verbatim}

    In the table below are the documents of the giant connected component
represented in the interactive graph above.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_86_0.png}\end{center}
        \caption{}
        \label{}
    \end{figure}

    Now we want to apply the same logic with the papers in-conferences. So
we will focus on the analysis of the first graph that uses the
intersection between the referenced documents of each two by two papers.
The length of the intersection set had to be more than 4.

So if more than 4 documents are referenced by two distinct documents
than the latter are more likely to be about the same subect.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_88_0.png}\end{center}
        \caption{}
        \label{}
    \end{figure}

    When applying this kind of clustering we find that most of the clusters
contains almost the same authors with different papers, which is logic
because if they are specialists in one topic and write several papers on
it, they might reuse some references in several documents.
